Rotational Spectral Unmixing of Exoplanets:\\Degeneracies between Surface Colors and Geography

### 2017-03-07 paper/main.tex
updating text

We were motivated by this unsatisfactory result to revisit their analysis. 
Specifically, the updates include:
\begin{itemize}
\item Revisiting the formulation and point out the inherit degeneracy between surface spectra and geography. 
\item Introducing regularization terms to enforce smooth longitudinal maps 
\end{itemize}

@@

The aim of this paper is to present how to constrain the albedo spectra of representative surface types from multi-band light curves. 
In order to facilitate the discussions about the procedures in the following sections, in this section we shall introduce a dataset to be used for demonstrations.

@@

The bottom panels of Figure \ref{fig:mockdata} display the examples of diurnal light curves in 4 photometric bands, represented in terms of {\it apparent albedo}. 
Apparent albedo is the planetary intensity normalized by that of a loss-less Lambert sphere with the same radius and at the same phase \citep{Qiu2003, Seager2010}; in this paper we will use apparent albedo unless otherwise noted. 
The apparent albedo is straightforwardly obtained from the observed planetary intensity, if and only if the planetary radius and the observational geometry are completely known: latitude/longitude of the sub-stellar/sub-observer points, and the distance between the star and the planet. 
\memoYF{Later in the paper, mention that the uncertainty in radius estimation severely affects the constraints. }
%(We will discuss the effect of this assumptions later in Section ??.)

@@

This (hyper-)plane can be identified through Principal Component Analysis (PCA) \citep{Cowan2009,Cowan2011}, as PCA extracts the major, mutually-orthogonal axes along which the scatter among the data points are significant. 
%Note that the number of principal components In other words, the number of spectrally distinct major surface types ($=K$) is the number of dominant principal components (PCs) plus 1 in theory \citep{Cowan2011}. 
% used Principal Component Analysis (PCA) to 
% The original data, however, have 4 photometric bands (i.e., $J=4$), and $J$-the dimensional space is inconvenient for visualization. 
%We thus start by performing principal component analysis (PCA). 
From the mock light curves prepared in Section \ref{s:mockdata}, we extract two dominant Principal Components (PCs) through PCA with other components having virtually zero contributions, consistent with the 3 input surface types.  
The PCs extracted from the light curves at $\alpha = 90^{\circ }$ (olive lines in Figure \ref{fig:mockdata}) are presented in the upper panel of Figure \ref{fig:trajectory}. 
In the following, we adopt these two PCs as the axes of the PC plane. \memoYF{Comment from Nick: This is true by definition?}
Although the PCs from other light curves with different phase angles are not necessarily same as the one shown in the upper panel of Figure \ref{fig:trajectory}, the PC plane itself should be the same plane (under the noiseless condition), because we assume the same 3 surface types. Thus it is ok to use these PCs to discuss other light curves as well.

The lower panel of Figure \ref{fig:trajectory} shows the trajectories of the mock light curves with varying phase angle on the plane defined by PC 1 and PC 2. 
The excursion of the trajectory is larger for the light curve at a larger phase angle, i.e., at crescent phase, and shrinks as the phase angle decreases because a larger surface area is averaged together.  
The points in the figure indicate the input albedo spectra of the three surface types. 
As described above, the trajectories are always inside the triangle defined by these points. 
%Because the apparent albedo at any time of the observations is a linear combination of input surface colors (Equation \ref{eq:tilde_d_f_ast_s}) under the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), the trajectories should always be inside the triangle defined by these points.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \begin{center}
\includegraphics[width=\hsize]{mockdata_PCplane.pdf}
    \end{center}
    \caption{Upper panel: principal components (PCs) of the light curves at $\alpha = 90^{\circ }$ shown in olive lines in Figure \ref{fig:mockdata}. Lower panel: trajectories of the 4-band light curves on the PC plane set by 2 PCs shown in the upper panel, in the case of $\alpha = 135^{\circ }$ (indigo thick line), $\alpha = 90^{\circ }$ (olive line), and $\alpha = 45^{\circ }$ (gold thin line). Points indicate the input albedo spectra of ocean (blue circle), soil (red square), and vegetation (green triangle) on the PC plane. Points in the gray region violate the condition (\ref{eq:tilde_s_range}); specifically, the left, right, and bottom boundaries are set by $s_{k,4}> 0$, $s_{k,1}> 0$, and $s_{k,3}> 0$, respectively. 
Dotted lines are random triangle that could be solutions (see text). }
    \label{fig:trajectory}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Additional constraint comes from the condition (\ref{eq:tilde_s_range}). 
In Figure \ref{fig:trajectory} any points in the shadowed region are rejected based on this condition: specifically, the left, right, and bottom boundaries are set by $s_{k,4}> 0$, $s_{k,1}> 0$, and $s_{k,3}> 0$. 
While in this particular example, the permitted region happen to have a triangle shape, the shapes can have a different shape depending on the relative configuration of the boundariese\memoYF{In more detail?}. 
Although these limits the region where $\{ {\bf s}_k \}$ can exist,   
this is in general not sufficient to result in a unique solution.

In Figure \ref{fig:trajectory} we show two random triangles that enclose the trajectory; these triangles satisfy the conditions, so do many others. 
One can have spectrally interesting surface spectra and boring geography (small longitudinal variation in area fractions) {\it or} one can have boring surface spectra (closer to flat) and interesting geography. \memoYF{Need to reconsider this phrasing. }
Therefore, predicting ${\bf \fast }$ and ${\bf s}$ from ${\bf d}$ is clearly a ill-posed problem.

A formally equivalent degeneracy is found in Equation (\ref{eq:d_f_s}) coupled with the conditions (\ref{eq:f_range})-(\ref{eq:s_range}). 
Essentially, the term $\sum _k f_{lk} s_{kj}$ represents the average albedo spectra of $l$-th slice and $W_{il}$ is the matrix that convolve them into the light curves. 
But again, we can have different sets of $\{ {\bf s}_k \}$ that enclose the averaged albedo spectra of longitudinal slices {\it and} are located in the range of condition (\ref{eq:s_range}), any of which can make up the given averaged albedo spectra with the associated $f_{lk}$. 
For disk-integrated light curves, the color excursions are more muted than the intrinsic longitudinal map, so the degeneracy becomes severer.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{``Best Guess'' of the Surface Types}
\label{ss:constraining}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_45deg_3types_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
%\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_lon_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types2_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage} 
    \caption{}%Left: Probability of surface albedo spectra estimated from colors of longitudinal slices, without regularization term (``flat'' prior?). Middle: on the assumption of gaussian process with correlation length $30^{\circ }$. Right panel: correlation length $40^{\circ }$. }
\label{fig:noreg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In order to see this more quantitatively, we make a grid on the PC plane  with an interval of 0.2 and consider all combinations of three grid points. 
Then, we assumed that all of the combinations that enclose the trajectory are equally likely to be a solution, and found the marginalized probability of having a surface type at each location of the PC plane. 
Specifically, we consider the following Bayesean expression and attempt to find the posterior distribution:
%%%
\begin{equation}
P_{\rm posterior} ( s_{kn}' | d_{ij} ) = P ( d_{ij} | s_{kn}' ) \cdot P_{\rm prior} (  s_{kn}' ) 
\end{equation}
%%%
where $s_{kn}'$ represents the coordinates of the $k$-th surface type on the PC plane ($n$ is either 1 or 2, corresponding to PC 1 and PC 2): $ s_{kn}' = s_{kj} V_{jn} $. 
Assuming that any facet has the same prior probability per area of being a surface type, the prior for each $k$ is:
%%%
\begin{eqnarray}
P_{\rm prior} (s_{k1}', s_{k2}' ) \, ds_{k1}' ds_{k2}' &=& \frac{ds_{k1}'  ds_{k2}'}{\iint _{\rm allowed} d^2{\bf s}_k'}. 
\end{eqnarray}
%%%
On the other hand, the likelihood is 
%%%
\begin{equation}
P ( d_{ij} | s_{kn}' ) \propto \left\{
\begin{array}{ll}
1 \;\;\; & \mbox{if satisfies conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum})} \\
0 \;\;\; & \mbox{otherwise}
\end{array}
\right.
\end{equation}
%%%
Although the graphical meaning of conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) is simply to enclose the data, the judgement is numerically made as follows. 
We first find $\fast $ that satisfies condition (\ref{eq:tilde_f_sum}) from
%%%
\begin{equation}
\begin{pmatrix}
d'_{11} & d'_{12} & 1 \\
... & & 1 \\
d'_{I1} & d'_{I2} & 1 
\end{pmatrix}
= 
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
\begin{pmatrix}
s'_{11} & s'_{12} & 1 \\
s'_{21} & s'_{22} & 1 \\
s'_{31} & s'_{32} & 1 
\end{pmatrix}
\end{equation}
%%%
or,
\begin{equation}
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
=
\begin{pmatrix}
d'_{11} & d'_{12} & 1 \\
... & & 1 \\
d'_{I1} & d'_{I2} & 1 
\end{pmatrix}
\begin{pmatrix}
s'_{11} & s'_{12} & 1 \\
s'_{21} & s'_{22} & 1 \\
s'_{31} & s'_{32} & 1 
\end{pmatrix}^{-1}
\end{equation}
%%%
Note that the last matrix is always a $K\times K$ matrix, and thus inverse is usually found. 
Then. we ask if the resultant $\fast_{ik}$ are within the range of (\ref{eq:tilde_f_range}).

Finally, we marginalize the posterior distribution, i.e.,
%%%
\begin{equation} 
P_{\rm posterior} ( {\bf s}_1) = \iint P_{\rm posterior} ( s_{kn}' | d_{ij} ) d^2 {\bf s}_2 d^2 {\bf s}_3
\end{equation}
%%%
This is because of the symmetry of ${\bf s}_1$, ${\bf s}_2$, and ${\bf s}_3$. 
$P_{\rm posterior} ( {\bf s}_1) $ is normalized so that it sums up to unity.

The resultant probability distribution is shown in Figure \ref{fig:noreg}. 
As expected, we find a relatively definite peak near the point of ocean. 
On the other hand, the peak that corresponds to soil is a blur, and it is almost not constrained. 
The vegetation in the bottom right corner is moderately constrained. 
We also have to note that

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Additional Constraints}
\label{ss:regularization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A possible constraining condition we might impose is the correlation length of the light curves, or of the map. 
Smooth time or longitudinal variation of covering area fraction with correlation lengths may be imposed as Gaussian Process \citep{Rasmussen2005}, which means the any finite number of the covering fractions have a joint Gaussian distribution. 
Note that the actual covering fraction of surface types do not necessarily obey gaussian process. 
Therefore, the prior knowledge may rather keep the ``solution'' away from the ground truth.

%%%
\begin{eqnarray}
P_{\rm posterior} ( s_{kn}' | d_{ij} ) &=& P ( d_{ij} | s_{kn}', f_{lk} ) \\
&& \times \, P_{\rm prior} (  s_{kn}' ) P_{\rm prior} (  f_{lk} ) \\
P ( d_{ij} | s_{kn}', f_{lk} ) &=& \delta( f_{lk} - f'_{lk} )
\end{eqnarray}
%%%

%%%
\begin{equation}
P_{\rm prior} (  f_{lk} ) \propto \frac{1}{\sqrt{| \Sigma^{(k)} |}}  \exp\left[ - (f^{T})_{kl} ( \Sigma ^{(k)} )^{-1}_{ll'} f_{l'k} \right]
\end{equation}
%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh]
    \begin{center}
	\includegraphics[width=\hsize]{raddata_2_PCplane_noreg.pdf}
    \end{center}
    \caption{a}
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We were motivated by this unsatisfactory result to revisit their analysis. 
Specifically, the updates are (1) Revisiting the formulation and point out the inherit degeneracy between surface spectra and geography, and (2) Introducing regularization terms to enforce smooth longitudinal maps. 
We limit our discussions to the analysis of noise-free data. Clearly, noise in realistic observations of exoplanets are expected to be substantially larger, and the effect of such observations will be discussed elsewhere.

@@

%The aim of this paper is to present how to constrain the albedo spectra of representative surface types from multi-band light curves. 
In order to facilitate the discussions in the following sections about the procedures to constrain the albedo spectra of representative surface types from multi-band light curves, we shall introduce a dataset to be used for demonstrations in this section.

@@

The bottom panels of Figure \ref{fig:mockdata} display the examples of diurnal light curves in 4 photometric bands, represented in terms of {\it apparent albedo}. 
Apparent albedo is the planetary intensity normalized by that of a loss-less Lambert sphere with the same radius and at the same phase \citep{Qiu2003, Seager2010}; in this paper we will use apparent albedo unless otherwise noted. 
The apparent albedo is straightforwardly obtained from the observed planetary intensity, if and only if the planetary radius and the observational geometry are completely known: latitude/longitude of the sub-stellar/sub-observer points, and the distance between the star and the planet. 
%\memoYF{Later in the paper, mention that the uncertainty in radius estimation severely affects the constraints. }
%(We will discuss the effect of this assumptions later in Section ??.)

@@

This (hyper-)plane can be identified through Principal Component Analysis (PCA) \citep{Cowan2009,Cowan2011}, as PCA extracts the major, mutually-orthogonal axes along which the scatter among the data points are significant. 
Consequently, the number of major principal components (PCs) are $K-1$ \citep{Cowan2011}. 
%Note that the number of principal components In other words, the number of spectrally distinct major surface types ($=K$) is the number of dominant principal components (PCs) plus 1 in theory \citep{Cowan2011}. 
% used Principal Component Analysis (PCA) to 
% The original data, however, have 4 photometric bands (i.e., $J=4$), and $J$-the dimensional space is inconvenient for visualization. 
%We thus start by performing principal component analysis (PCA). 
From the mock light curves prepared in Section \ref{s:mockdata}, we extract two dominant PCs through PCA with other components having virtually zero contributions, consistent with the 3 input surface types.  
The PCs extracted from the light curves at $\alpha = 90^{\circ }$ (olive lines in Figure \ref{fig:mockdata}) are presented in the upper panel of Figure \ref{fig:trajectory}. 
In the following, we adopt these two PCs as the axes of the PC plane. 
\memoYF{Comment from Nick: This is true by definition?}
For the later use, we denote these PCs by $V_{nj}$ where $n=1$ or $2$  corresponding to the 1st and 2nd PC. 
Although the PCs from other light curves with different phase angles are not necessarily same as the one shown in the upper panel of Figure \ref{fig:trajectory}, the PC plane itself should be the same plane (under the noiseless condition), because we assume the same 3 surface types. 
Thus in this section we keep using is PCs to discuss other light curves as well.

The lower panel of Figure \ref{fig:trajectory} shows the trajectories of the mock light curves with varying phase angle on the plane defined by PC 1 and PC 2. 
For the later use, these trajectories on the PC plane is denoted by $U_{in}$, in other words 
%%%
\begin{equation}
d_{ij} = \sum_n U_{in} V_{nj} + \bar d_j
\end{equation}
%%%
where $\bar d_j$ is the offset of the PC plane, which is in our case the time average of apparent albedo in the case of $\alpha = 90^{\circ }$. 
The excursion of the trajectory is larger for the light curve at a larger phase angle, i.e., at crescent phase, and shrinks as the phase angle decreases because a larger surface area is averaged together.  
The points in the figure indicate the input albedo spectra of the three surface types. 
As described above, the trajectories are always inside the triangle defined by these points. 
%Because the apparent albedo at any time of the observations is a linear combination of input surface colors (Equation \ref{eq:tilde_d_f_ast_s}) under the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), the trajectories should always be inside the triangle defined by these points.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \begin{center}
\includegraphics[width=\hsize]{mockdata_PCplane.pdf}
    \end{center}
    \caption{Upper panel: principal components (PCs) of the light curves at $\alpha = 90^{\circ }$ shown in olive lines in Figure \ref{fig:mockdata}. Lower panel: trajectories of the 4-band light curves on the PC plane set by 2 PCs shown in the upper panel, in the case of $\alpha = 135^{\circ }$ (indigo thick line), $\alpha = 90^{\circ }$ (olive line), and $\alpha = 45^{\circ }$ (gold thin line). Points indicate the input albedo spectra of ocean (blue circle), soil (red square), and vegetation (green triangle) on the PC plane. Points in the gray region violate the condition (\ref{eq:tilde_s_range}); specifically, the left, right, and bottom boundaries are set by $s_{k,4} > 0$, $s_{k,1} > 0$, and $s_{k,3}> 0$, respectively. 
Dotted lines are random triangle that could be solutions (see text). }
    \label{fig:trajectory}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Additional constraint comes from the condition (\ref{eq:tilde_s_range}). 
In Figure \ref{fig:trajectory} any points in the shadowed region are rejected based on this condition: specifically, the left, right, and bottom boundaries are set by $s_{k,4}> 0$, $s_{k,1}> 0$, and $s_{k,3}> 0$. 
While in this particular example the permitted region happen to have a triangle shape, the shapes can have a different shape depending on the relative configuration of the boundariese\memoYF{In more detail?}. 
Although these limits the region where $\{ {\bf s}_k \}$ can exist,   
this is in general not sufficient to result in a unique solution.

In Figure \ref{fig:trajectory} we show two random triangles that enclose the trajectory; these triangles satisfy the conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}), so do many others. 
One can have spectrally interesting surface spectra and boring geography (small longitudinal variation in area fractions) {\it or} one can have boring surface spectra (closer to flat) and interesting geography. \memoYF{Need to reconsider this phrasing. }
Therefore, predicting ${\bf \fast }$ and ${\bf s}$ from ${\bf d}$ is clearly a ill-posed problem.

A formally equivalent degeneracy is found in Equation (\ref{eq:d_f_s}) coupled with the conditions (\ref{eq:f_range})-(\ref{eq:s_range}). 
Essentially, the term $\sum _k f_{lk} s_{kj}$ represents the average albedo spectra of $l$-th slice and $W_{il}$ is the matrix that convolve them into the light curves. 
Suppose idealistically that we can retrieve the average spectra of longitudinal slices from the light curves through $W_{il}$; the trajectory in the panel C of Figure \ref{fig:trajectory} represents the color variation along the longitude.  
Again, we can have different sets of $\{ {\bf s}_k \}$ that enclose the averaged albedo spectra of longitudinal slices {\it and} are located in the range of condition (\ref{eq:s_range}), any of which can make up the given averaged albedo spectra with the associated $f_{lk}$. 
Nevertheless, the excursions of the longitudinal colors are more dramatic than the disk-integrated colors of the light curves, thus the possible solutions are somewhat more restricted. 
% For disk-integrated light curves, the color excursions are more muted than the intrinsic longitudinal map, so the degeneracy becomes severer.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{``Best Guess'' of the Surface Types}
\label{ss:guess}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbh!]
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
    \end{center}
     \end{minipage}   
    \begin{minipage}{0.33\hsize}
    \begin{center}
%\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_noreg.pdf}
\includegraphics[width=\hsize]{mockdata_135deg_3types_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{IGBP_lon_noreg.pdf}
    \end{center}
     \end{minipage}
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types2_t360_lc_noreg.pdf}
    \end{center}
     \end{minipage} 
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_reg_l30deg.pdf}
    \end{center}
     \end{minipage}      
   \begin{minipage}{0.33\hsize}
    \begin{center}
\includegraphics[width=\hsize]{mockdata_90deg_3types_t12_lc_reg_l40deg.pdf}
    \end{center}
     \end{minipage}
    \caption{(From top-left) A: Probability of surface albedo spectra estimated from the light curves of $\alpha = 90^{\circ }$ in the procedure described in Section \ref{ss:guess}. B: same as A but from the light curves of  $\alpha = 135^{\circ }$. C: same as C but from the colors of longitudinal slices.  D: same as A but from the light curves where the geographical map of soil and vegetation are switched. E: same as A but with regularization with $\lambda = 30^{\circ }$. same as A but with regularization with $\lambda = 40^{\circ }$. }
\label{fig:noreg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In order to see this more quantitatively, we make a grid on the PC plane  with an interval of 0.2 and consider all combinations of three grid points. 
Then, we assumed that all of the combinations that enclose the trajectory are equally likely to be a solution, and found the marginalized probability of having a surface type at each location of the PC plane. 
Specifically, we consider the following Bayesean expression and attempt to find the posterior distribution:
%%%
\begin{equation}
P_{\rm posterior} ( t_{kn} | U_{in} ) = P ( U_{in} | t_{kn} ) \cdot P_{\rm prior} (  t_{kn} ) 
\end{equation}
%%%
where $t_{kn}$ represents the coordinates of the $k$-th surface type on the PC plane ($n$ is either 1 or 2, corresponding to PC 1 and PC 2), i.e., 
%%%
\begin{equation}
s_{kj} = \sum_n t_{kn} V_{nj} + \bar d_j . 
\end{equation}
Assuming that any facet has the same prior probability per area of being a surface type, the prior for each $k$ is:
%%%
\begin{eqnarray}
&& P_{\rm prior} (t_{k1}, t_{k2} ) \, dt_{k1} dt_{k2} \\
&& = \left\{
\begin{array}{ll}
\displaystyle \frac{dt_{k1}  dt_{k2}}{\iint _{\rm allowed} d^2{\bf t}_k} & \;\;\;\;\;\; \mbox{if satisfies (\ref{eq:tilde_s_range})} \\
0 & \;\;\;\;\;\; \mbox{otherwise}
\end{array}
\right.
\end{eqnarray}
%%%
On the other hand, the likelihood is 
%%%
\begin{eqnarray}
&& P ( U_{in} | t_{kn} ) \\
&& \propto \left\{
\begin{array}{ll}
1 \;\;\; & \;\;\;\;\;\; \mbox{if satisfies (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_range})} \\
0 \;\;\; & \;\;\;\;\;\; \mbox{otherwise}
\end{array}
\right.
\end{eqnarray}
%%%
Although the graphical meaning of conditions (\ref{eq:tilde_f_range}) and (\ref{eq:tilde_f_sum}) is simply to enclose the data, the judgement is numerically made as follows. 
We first find $\fast $ that satisfies condition (\ref{eq:tilde_f_sum}) from
%%%
\begin{equation}
\begin{pmatrix}
U_{11} & U_{12} & 1 \\
... & & 1 \\
U_{I1} & U_{I2} & 1 
\end{pmatrix}
= 
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
\begin{pmatrix}
t_{11} & t_{12} & 1 \\
t_{21} & t_{22} & 1 \\
t_{31} & t_{32} & 1 
\end{pmatrix}
\end{equation}
%%%
or,
\begin{equation}
\begin{pmatrix}
\fast_{11} & \fast_{12} & \fast_{13}  \\
... & \\
\fast_{I1} & \fast_{I2} & \fast_{I3}
\end{pmatrix}
=
\begin{pmatrix}
U_{11} & U_{12} & 1 \\
... & & 1 \\
U_{I1} & U_{I2} & 1 
\end{pmatrix}
\begin{pmatrix}
t_{11} & t_{12} & 1 \\
t_{21} & t_{22} & 1 \\
t_{31} & t_{32} & 1 
\end{pmatrix}^{-1}
\label{eq:f=ds-1}
\end{equation}
%%%
Note that the last matrix is a $K\times K$ matrix, and thus inverse is found if $\{ {\bf t}_k \}$ forms a triangle (rather than a straight line or a single dot, in which case $\{ {\bf t}_k \}$ is not a solution anyways).
Then. we ask if the resultant $\fast_{ik}$ are within the range of (\ref{eq:tilde_f_range}).

Finally, we marginalize the posterior distribution, i.e.,
%%%
\begin{equation} 
P_{\rm posterior} ( {\bf t}_1) = \iint P_{\rm posterior} ( t_{kn} | d_{ij} ) d^2 {\bf t}_2 d^2 {\bf t}_3
\end{equation}
%%%
given the symmetry of ${\bf t}_1$, ${\bf t}_2$, and ${\bf t}_3$. 
$P_{\rm posterior} ( {\bf t}_1) $ is normalized so that it sums up to unity.

The resultant probability distribution is shown in Figure \ref{fig:noreg}. 
As expected, we find a relatively definite peak near the point of ocean. 
On the other hand, the peak that corresponds to soil is a blur, and it is almost not constrained. 
The vegetation in the bottom right corner is moderately constrained.

We iterate that in this framework the success of the guess critically depends on the relative configuration of the trajectory and allowed region. 
In panel D of Figure \ref{fig:noreg}, we show another example where we switched the geographical map of soil and vegetation; as a result, the trajectory moves toward soil rather than vegetation. However, because the allowed region near the soil is large, soil is still not sufficiently constrained.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Additional Assumptions}
\label{ss:regularization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A possible constraining assumption we might make is the correlation length of the area fraction as a function of time, or as a function of longitude.  
Smooth time or longitudinal variation with correlation lengths may be imposed as Gaussian Process \citep{Rasmussen2005}, which means the any finite number of the covering fractions have a joint Gaussian distribution. 
In our case, the prior for $f_{lk}$ for each $k$ is then:
%%%
\begin{eqnarray}
&& P_{\rm prior} (  \fast_{1k}, \fast_{2k}, ..., \fast_{Ik} ) \\
&& \propto \frac{1}{\sqrt{| \tilde \Sigma^{(k)} |}}  \exp\left[ - (\fast^{T})_{ki} ( \tilde \Sigma ^{(k)} )^{-1}_{ii'} \fast_{i'k} \right]
\end{eqnarray}
%%%
where
%%%
\begin{equation}
\tilde \Sigma _{ll'} = \lambda \exp \left( - \frac{|t_i- t_i'|^2}{2 \Delta_ t^2} \right)
\end{equation}
%%%
or
%%%
\begin{eqnarray}
&& P_{\rm prior} (  f_{1k}, f_{2k}, ..., f_{Lk} ) \\
&& \propto \frac{1}{\sqrt{| \Sigma^{(k)} |}}  \exp\left[ - (f^{T})_{kl} ( \Sigma ^{(k)} )^{-1}_{ll'} f_{l'k} \right]
\end{eqnarray}
%%%
where
%%%
\begin{equation}
\Sigma _{ll'} = \lambda \exp \left( - \frac{|\phi_l- \phi_{l'}|^2}{2 \Delta _{\phi }^2} \right). 
\end{equation}
%%%

The posterior will be
%%%
\begin{eqnarray}
P_{\rm posterior} ( t_{kn} | d_{ij} ) &=& P ( d_{ij} | t_{kn}, f_{lk} ) \\
&& \times \, P_{\rm prior} (  t_{kn} ) P_{\rm prior} (  f_{lk} ) \\
P ( d_{ij} | t_{kn}, f_{lk} ) &=& \delta( f_{lk} - f_{lk}^{(0)} (t_{kn}) )
\end{eqnarray}
%%%
where $f_{lk}^{(0)}$ is the solution of Equation (\ref{eq:f=ds-1}). 
\memoYF{Very acrobatic. Am I doing right? I am not sure.}

Panel E and F of Figure \ref{fig:noreg} are the results where $\Delta_ t = 2$ hr and $2.7$ hr, respectively. 
In this case, we find the higher peak of albedo spectra of ocean closer to the ``answer'', while the vegetation is identified far off the ``answer''; vegetation is intermediate.

We must take caution here that the actual covering fraction of surface types do not necessarily obey gaussian process. 
In this particular case, it appears that the dominant surface types are better approximated by gaussian processes.... \memoYF{??}

\memoYF{Have to do with regularization on the maps}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EPOXI data}
\label{s:EPOXI}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbh!]
    \begin{center}
	\includegraphics[width=\hsize]{raddata_2_PCplane_noreg.pdf}
    \end{center}
    \caption{}
\label{fig:EPOXI}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we apply our analysis procedure to the observed multi-band light curves of the Earth observed by EPOXI mission. 
Specifically we use the same datasets as \citet{Cowan2013}, i.e., the diurnal light curves observed in June of 2008. 
Using two dominant PCs we extracted, we performed the same analysis described in the previous section

Figure \ref{fig:EPOXI} shows the trajectory of the light curve projected onto the PC plane, with the grayed region being the forbidden region. 
While the permitted region appears to be close to a square, it is in fact a heptagon, bounded by 7 inequalities. Overall, the upper boundaries come from the requirement of $s_{kj}<1$ while the lower boundaries comes from the requirement of $s_{kj}<0$.

In this case, the trajectory of planetary light curves is away from the boundary because the color variations of the real Earth are muted by the cloud cover \memoYF{Right?}. 
Consequently, we do not see three peaks but a spectrum of the potential. 
Nevertheless, we are likely to find one surface type near the upper right part of the allowed region.

\memoYF{Draw a spectra at the upper left!!}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{s:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

So far we have assumed that the geometrical parameters are completely known, including the planetary orbit, planetary spin axis and the planetary radius. 
In this section we discuss the influence of incomplete knowledge of these parameters.

The decomposition into $\fast $ and $s$ does not require the information of spin axis or spin period. 
What matters most in this case is the uncertainty in the normalization of apparent albedo. 
In reality, direct imaging observations provide the intensity of the 
planetary light and the intensity of the stellar light. 
In order to convert this to the unit of albedo, we need to know the orbital distance from the star, phase angle, and the planetary radius. 
If there is uncertainties in these measurements, the constraints originating from the upper boundary for albedo, $s_{jk} < 1$ may be unconstrained. 
For example, in the case of EPOXI data, about a half of the boundaries would be uncertain if the normalization of albedo is uncertain. 
Note that this is not always the case; in the case of our mockdata presented in Sections \ref{s:mockdata} and \ref{s:frame}, by coincidence \memoYF{?} all of the boundaries come from non-negativity of albedo, $s_{jk} > 0$. 
Thus, depending on the nature of the planetary surface, there is a possibility that the planetary radius is now strongly affected by the unknown planetary radius, orbital distance, or orbital phase. \memoYF{Really?????}

If the normalization of apparent albedo is not known, we do not have the boundaries

\memoYF{Instability of inverse problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{s:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this paper, we revisited the framework to estimate the albedo spectra of major surface types of exoplanets and their distribution from disk-integrated multi-band photometric data. 
We pointed out the inherit degeneracy which makes it impossible to find unique solutions. 
Under some assumptions, constraints may be found about the likely solutions, due to the nature of albedo which is between 0 and 1. 
We demonstrate how such constraints may be obtained using a simplified toy model of the Earth.

The constraints can be
